---
title: "Research Metrics Report: Overhaul Draft"
subtitle: |
  John G. Shedd Aquarium
output:
  html_document:
    toc: true
    toc_depth: 2
date: "`r format(Sys.time(), '%B %d, %Y')`"
params:
  cutoff: !r Sys.Date()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 4,
  fig.width = 8
)

cutoff <- as.Date(params$cutoff)
```

```{r libraries}
library(readxl)
library(tidyverse)
library(lubridate)
library(janitor)
library(ggthemes)
library(RColorBrewer)
library(stringdist)

theme_custom <- function() {
  theme_few() +
    theme(
      legend.title = element_blank(),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.justification = "center",
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 11),
      strip.text = element_text(size = 13),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
    )
}

quarter_levels <- paste0("Q", 1:4)

clean_number <- function(x) {
  as.numeric(str_extract(as.character(x), "-?[0-9]+\\.?[0-9]*"))
}

first_non_missing_date <- function(...) {
  dates <- list(...)
  out <- as.Date(rep(NA, length(dates[[1]])))
  for (date_col in dates) {
    date_col <- as.Date(date_col)
    out <- coalesce(out, date_col)
  }
  out
}

standardize_category <- function(category) {
  category_norm <- str_squish(str_replace_all(category, "\u00a0", " "))

  case_when(
    str_detect(category_norm, regex("citizen/stakeholder|community engagement|community / stakeholder engagement|capacity sharing|capacity", ignore_case = TRUE)) ~ "Capacity Sharing & Training",
    str_detect(category_norm, regex("media", ignore_case = TRUE)) ~ "Media opportunity",
    str_detect(category_norm, regex("field research|field days|field", ignore_case = TRUE)) ~ "Field research",
    str_detect(category_norm, regex("publication|publications", ignore_case = TRUE)) ~ "Publication",
    str_detect(category_norm, regex("presentation|academic|conference|active meeting|meeting|tour", ignore_case = TRUE)) ~ "Presentation",
    TRUE ~ category_norm
  )
}

standardize_presentation_type <- function(type) {
  type_norm <- str_squish(type)

  case_when(
    type_norm %in% c("Panel discussion") ~ "Panel",
    TRUE ~ type_norm
  )
}

metric_summary <- function(data, category, value = c("n", "quantity"), pre2019_category = category) {
  value <- match.arg(value)

  current <- data %>%
    filter(Category == category) %>%
    group_by(Year, Quarter) %>%
    summarise(
      Quantity = if (value == "n") n() else sum(Quantity, na.rm = TRUE),
      .groups = "drop"
    )

  historical <- pre2019 %>%
    filter(Category == pre2019_category) %>%
    transmute(Year, Quarter, Quantity = clean_number(Quantity))

  bind_rows(current, historical) %>%
    mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
    complete(
      Year = full_seq(range(Year, na.rm = TRUE), 1),
      Quarter = factor(quarter_levels, levels = quarter_levels),
      fill = list(Quantity = 0)
    ) %>%
    arrange(Year, Quarter)
}

plot_quarterly_bars <- function(summary_data, title, label_nudge) {
  ggplot(summary_data, aes(x = Quarter, y = Quantity, fill = factor(Year))) +
    facet_grid(~ Year) +
    geom_col(alpha = 0.75) +
    geom_text(
      aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
      color = "white",
      nudge_y = label_nudge,
      size = 3
    ) +
    theme_custom() +
    theme(legend.position = "none", axis.text.x = element_text(size = 7)) +
    labs(x = "", y = "", title = title)
}

plot_annual_bars <- function(summary_data, title, label_nudge) {
  year_breaks <- seq(min(summary_data$Year, na.rm = TRUE), max(summary_data$Year, na.rm = TRUE), by = 1)

  summary_data %>%
    group_by(Year) %>%
    summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop") %>%
    ggplot(aes(x = Year, y = Quantity, fill = factor(Year))) +
    geom_col(alpha = 0.75) +
    geom_text(
      aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
      color = "white",
      nudge_y = label_nudge,
      size = 3
    ) +
    theme_custom() +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = year_breaks) +
    labs(x = "", y = "", title = title)
}

dedupe_references <- function(data, reference_col = Citation, threshold = 0.10) {
  data <- data %>%
    filter(!is.na({{ reference_col }}), {{ reference_col }} != "")

  if (nrow(data) < 2) {
    return(data)
  }

  groups <- data %>%
    pull({{ reference_col }}) %>%
    stringdistmatrix(method = "jw") %>%
    as.dist() %>%
    hclust(method = "average") %>%
    cutree(h = threshold)

  data[!duplicated(groups), ]
}
```

# Import And Normalize Data

```{r import-data}
pre2019 <- read_xlsx("data/pre2019_summary.xlsx") %>%
  mutate(
    Year = as.integer(Year),
    Quarter = factor(Quarter, levels = quarter_levels),
    Quantity = clean_number(Quantity)
  )

metrics_2019_2022 <- read_xlsx("data/research_metrics_2019-2022.xlsx") %>%
  transmute(
    Source = "2019-2022 Microsoft Form",
    ID = as.character(ID),
    Email,
    Name,
    Department,
    Date = as.Date(`Date of event`),
    RawCategory = Category,
    Category = standardize_category(Category),
    MediaOpportunityType = NA_character_,
    PresentationType = NA_character_,
    CapacitySharingType = NA_character_,
    Citation,
    Link,
    Effort = clean_number(Effort),
    Quantity = clean_number(Quantity),
    FieldDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldPersonDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldConservationResearch = if_else(Category == "Field research", Quantity, NA_real_),
    FieldAnimalHealth = NA_real_,
    FieldAnimalCare = NA_real_,
    FieldOtherDepartments = NA_real_,
    Description,
    Program,
    Pillar
  )

metrics_2022_2024_raw <- read_xlsx(
  "data/research_metrics_2022-2024.xlsx",
  col_types = c(rep("guess", 8), "text", "text", rep("guess", 26))
)

metrics_2022_2024 <- metrics_2022_2024_raw %>%
  transmute(
    Source = "2022-2024 IT Form",
    ID = as.character(ID),
    Email,
    Name,
    Department = `Select your Animal Care & Science department`,
    SingleMulti = `Was this a single-day or multi-day engagement?`,
    DateEngagement = as.Date(`What was the date of the engagement?`),
    DateStart = as.Date(`What was the starting date of the engagement?`),
    DateEnd = as.Date(`What was the ending date of the engagement?`),
    DatePublication = as.Date(`Publication Date`),
    DateMedia = as.Date(`Media Opportunity Date`),
    RawCategory = `What was the engagement type?`,
    Category = standardize_category(RawCategory),
    MediaOpportunityType = NA_character_,
    PresentationType = NA_character_,
    CapacitySharingType = NA_character_,
    Citation,
    Link = case_when(
      Category == "Publication" ~ `Publication Link`,
      Category == "Media opportunity" ~ `Media Link`,
      TRUE ~ NA_character_
    ),
    Effort = clean_number(`Community / Stakeholder Engagement Effort`),
    Quantity = case_when(
      Category == "Field research" & SingleMulti == "Multi-day" ~ as.numeric(DateEnd - DateStart) + 1,
      Category == "Field research" & SingleMulti == "Single-day" ~ 1,
      Category == "Capacity Sharing & Training" ~ clean_number(`Community / Stakeholder Engagement Quantity`),
      TRUE ~ NA_real_
    ),
    FieldDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldPersonDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldConservationResearch = if_else(Category == "Field research", Quantity, NA_real_),
    FieldAnimalHealth = NA_real_,
    FieldAnimalCare = NA_real_,
    FieldOtherDepartments = NA_real_,
    Description = `Briefly describe, or enter any notes for, this engagement`,
    Program = `If the engagement is for a program, select it from the list below`,
    Pillar = NA_character_,
    Date = first_non_missing_date(DatePublication, DateMedia, DateEngagement, DateStart, as.Date(`Start time`))
  ) %>%
  select(-SingleMulti, -DateEngagement, -DateStart, -DateEnd, -DatePublication, -DateMedia)

new_metrics_file <- list.files(path = "data", pattern = "Research metrics new.*\\.xlsx$", full.names = TRUE) %>%
  sort() %>%
  tail(1)

metrics_2024_on <- read_xlsx(new_metrics_file) %>%
  transmute(
    Source = "2024-on Microsoft Form",
    ID = as.character(Id),
    Email,
    Name,
    Department,
    Date = as.Date(`Date of event`),
    RawCategory = Category,
    Category = standardize_category(Category),
    MediaOpportunityType = str_squish(`Type of media opportunity`),
    PresentationType = standardize_presentation_type(`Type of presentation`),
    CapacitySharingType = str_squish(`Type of capacity sharing & training`),
    Citation,
    Link,
    Effort = clean_number(coalesce(as.character(Effort), as.character(Effort1))),
    Quantity = clean_number(Quantity),
    FieldDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldConservationResearchHeadcount = clean_number(`Field Team: Number of members from Conservation Research`),
    FieldAnimalHealthHeadcount = clean_number(`Field Team: Number of members from Animal Health`),
    FieldAnimalCareHeadcount = clean_number(`Field Team: Number of members from Animal Care`),
    FieldOtherDepartmentsHeadcount = clean_number(`Field Team: Number of members from other Shedd Departments`),
    FieldTeamHeadcount = rowSums(
      across(ends_with("Headcount"), ~ replace_na(.x, 0)),
      na.rm = TRUE
    ),
    FieldConservationResearch = case_when(
      Category == "Field research" & FieldTeamHeadcount > 0 ~ Quantity * FieldConservationResearchHeadcount,
      Category == "Field research" ~ Quantity,
      TRUE ~ NA_real_
    ),
    FieldAnimalHealth = if_else(Category == "Field research", Quantity * replace_na(FieldAnimalHealthHeadcount, 0), NA_real_),
    FieldAnimalCare = if_else(Category == "Field research", Quantity * replace_na(FieldAnimalCareHeadcount, 0), NA_real_),
    FieldOtherDepartments = if_else(Category == "Field research", Quantity * replace_na(FieldOtherDepartmentsHeadcount, 0), NA_real_),
    FieldPersonDays = case_when(
      Category == "Field research" & FieldTeamHeadcount > 0 ~ Quantity * FieldTeamHeadcount,
      Category == "Field research" ~ Quantity,
      TRUE ~ NA_real_
    ),
    Description,
    Program,
    Pillar
  ) %>%
  select(-ends_with("Headcount"), -FieldTeamHeadcount)

metrics_all <- bind_rows(metrics_2019_2022, metrics_2022_2024, metrics_2024_on) %>%
  mutate(
    Year = year(Date),
    Quarter = factor(paste0("Q", quarter(Date)), levels = quarter_levels),
    Department = coalesce(Department, "Conservation Research")
  ) %>%
  mutate(
    DescriptionKey = str_squish(str_to_lower(coalesce(Description, ""))),
    CitationKey = str_squish(str_to_lower(coalesce(Citation, ""))),
    LinkKey = str_squish(str_to_lower(coalesce(Link, ""))),
    DuplicateKey = case_when(
      Category == "Publication" & CitationKey != "" ~ paste(Category, CitationKey),
      LinkKey != "" ~ paste(Category, Date, LinkKey),
      DescriptionKey != "" ~ paste(Category, Date, DescriptionKey),
      TRUE ~ paste(Category, Date, Name, Quantity)
    )
  ) %>%
  filter(!duplicated(DuplicateKey)) %>%
  select(
    -DuplicateKey,
    -DescriptionKey,
    -CitationKey,
    -LinkKey
  )

metrics <- metrics_all %>%
  filter(!is.na(Date), Date <= cutoff)
```

```{r import-audit}
metrics %>%
  count(Source, Category, sort = TRUE) %>%
  knitr::kable(caption = "Rows imported by source and standardized metric category")
```

## Working Assumptions

This overhaul standardizes historical metric names to the current reporting
names. `Citizen/stakeholder engagement`, `Community / Stakeholder Engagement`,
`Community Engagement`, and `Capacity Sharing & Training` are all reported as
`Capacity Sharing & Training`, with the reported quantity interpreted as number
of people engaged. `Conference or active meeting`, `Conference/Meeting`,
`Academic`, `Tour`, and other presentation-like labels are reported as
`Presentation`. `Media` is reported as `Media opportunity`, and old/new field
labels are reported as `Field research`.

For field research, older rows without team headcounts are carried forward as
both field days and Conservation Research person-days. Newer rows with
department headcounts use:

`field person-days = field days * total Shedd field team headcount`

# Publications & Citations

```{r citation-data}
citation_cache <- "data/citation_counts.txt"

citation_history <- read.delim(
  citation_cache,
  header = TRUE,
  sep = "\t",
  fill = TRUE,
  stringsAsFactors = FALSE
) %>%
  mutate(CacheRow = row_number()) %>%
  as_tibble() %>%
  clean_names() %>%
  mutate(date = as.Date(date))

if (!Sys.Date() %in% citation_history$date) {
  library(scholar)

  hc_authors <- read_tsv("data/authors.txt", show_col_types = FALSE) %>% drop_na()
  hc_oldrefs <- read_tsv("data/pre2019_publications.txt", col_names = "reference", show_col_types = FALSE)
  hc_newrefs <- metrics %>%
    filter(Category == "Publication") %>%
    select(reference = Citation)

  hc_references <- bind_rows(hc_oldrefs, hc_newrefs) %>%
    dedupe_references(reference)

  all_pubs <- hc_authors %>%
    mutate(all_pubs = map(scholar_id, get_publications)) %>%
    unnest(all_pubs)

  hc_pubs <- all_pubs %>%
    mutate(hc_affil = map_lgl(title, ~ any(agrepl(.x, hc_references$reference)))) %>%
    filter(hc_affil, !is.na(year)) %>%
    filter(!(duplicated(cid) & duplicated(tolower(title))))

  pub_cites <- hc_pubs %>%
    arrange(desc(cites)) %>%
    pull(cites)

  h.index <- tail(which(pub_cites >= seq_along(pub_cites)), 1)
  i10.index <- sum(pub_cites >= 10)

  todays_totals <- hc_pubs %>%
    summarise(
      date = Sys.Date(),
      total_pubs = nrow(hc_pubs),
      total_cites = sum(cites),
      h_index = h.index,
      i10_index = i10.index
    )

  bind_rows(citation_history, todays_totals) %>%
    select(-cache_row) %>%
    arrange(date) %>%
    write_tsv(citation_cache)

  citation_history <- read.delim(
    citation_cache,
    header = TRUE,
    sep = "\t",
    fill = TRUE,
    stringsAsFactors = FALSE
  ) %>%
    mutate(CacheRow = row_number()) %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(date = as.Date(date))
}

cites_raw <- citation_history %>%
  group_by(date) %>%
  slice_max(cache_row, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-cache_row) %>%
  filter(date <= cutoff)

latest <- cites_raw %>%
  arrange(date) %>%
  tail(1)

current_year <- year(cutoff)
current_quarter <- factor(paste0("Q", quarter(cutoff)), levels = quarter_levels)
period_fill_values <- c("Complete period" = "gray65", "Year to date" = "gray35", "Quarter to date" = "gray35")
```

## Current Publication Impact Snapshot

Latest cached citation pull: `r latest$date`

```{r citation-snapshot}
tibble(
  Metric = c("Publications", "Citations", "h-index", "i10-index"),
  `Current value` = c(
    scales::comma(latest$total_pubs),
    scales::comma(latest$total_cites),
    as.character(latest$h_index),
    as.character(latest$i10_index)
  )
) %>%
  knitr::kable(align = c("l", "r"))
```

## Publications

```{r publications}
nondupes <- metrics %>%
  filter(Category == "Publication") %>%
  dedupe_references(Citation)

pub_historical_annual <- read_lines("data/pre2019_publications.txt") %>%
  tibble(reference = .) %>%
  mutate(Year = as.integer(str_extract(reference, "\\((20[0-9]{2})\\)", group = 1))) %>%
  filter(!is.na(Year), Year <= 2018) %>%
  count(Year, name = "Quantity")

pub_quarterly_current <- nondupes %>%
  filter(Year >= 2019) %>%
  group_by(Year, Quarter) %>%
  summarise(Quantity = n(), .groups = "drop") %>%
  mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

pub_annual <- bind_rows(
  pub_historical_annual,
  pub_quarterly_current %>%
    group_by(Year) %>%
    summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")
) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    fill = list(Quantity = 0)
  )

pubsumm <- bind_rows(
  pub_historical_annual %>%
    mutate(Quarter = factor(NA_character_, levels = quarter_levels)),
  pub_quarterly_current
)

pub_recent_years <- pub_quarterly_current %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

pub_quarterly_recent <- pub_quarterly_current %>%
  filter(Year %in% pub_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

pub_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(pub_annual$Year), max(pub_annual$Year), by = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill_values[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Number of Scientific Publications")

pub_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill_values[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly Scientific Publications, Most Recent 5 Years")
```

## Citations

```{r citations}
citesumm <- cites_raw %>%
  mutate(
    Year = year(date),
    Quarter = factor(paste0("Q", quarter(date)), levels = quarter_levels)
  ) %>%
  group_by(Year, Quarter) %>%
  distinct(date, .keep_all = TRUE) %>%
  filter(date == max(date)) %>%
  ungroup() %>%
  arrange(Year, Quarter) %>%
  mutate(Quantity = total_cites - lag(total_cites)) %>%
  transmute(Year, Quarter, Quantity = replace_na(Quantity, 0)) %>%
  bind_rows(
    pre2019 %>%
      filter(Category == "Citations") %>%
      select(Year, Quarter, Quantity)
  ) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

cite_annual <- citesumm %>%
  group_by(Year) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")

cite_recent_years <- citesumm %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

cite_quarterly_recent <- citesumm %>%
  filter(Year %in% cite_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

cite_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(cite_annual$Year), max(cite_annual$Year), by = 1)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill_values[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Number of New Citations")

cite_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill_values[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly New Citations, Most Recent 5 Years")
```

# Media Opportunities

```{r media}
mediasumm <- metric_summary(
  metrics,
  category = "Media opportunity",
  value = "n",
  pre2019_category = "Media opportunity"
)

plot_quarterly_bars(mediasumm, "Number of Media Opportunities", -0.5)
plot_annual_bars(mediasumm, "Number of Media Opportunities", -0.5)
```

# Presentations

```{r presentations}
pressumm <- metric_summary(
  metrics,
  category = "Presentation",
  value = "n",
  pre2019_category = "Presentation"
)

plot_quarterly_bars(pressumm, "Number of Presentations", -0.75)
plot_annual_bars(pressumm, "Number of Presentations", -0.75)

presentation_type_quarterly <- metrics %>%
  filter(
    Category == "Presentation",
    !is.na(PresentationType),
    PresentationType != "",
    Year >= 2026
  ) %>%
  arrange(Year, Quarter) %>%
  mutate(
    PresentationType = fct_lump_min(PresentationType, min = 2, other_level = "Other single-entry"),
    YearQuarterLabel = if_else(
      Year == current_year & Quarter == current_quarter,
      paste(Year, Quarter, "(QTD)"),
      paste(Year, Quarter)
    ),
    YearQuarter = factor(YearQuarterLabel, levels = unique(YearQuarterLabel))
  ) %>%
  count(Year, Quarter, YearQuarter, PresentationType, name = "Quantity") %>%
  complete(
    YearQuarter,
    PresentationType,
    fill = list(Quantity = 0)
  ) %>%
  group_by(PresentationType) %>%
  mutate(TypeTotal = sum(Quantity, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(PresentationType = fct_reorder(PresentationType, TypeTotal))

ggplot(presentation_type_quarterly, aes(x = YearQuarter, y = PresentationType, fill = Quantity)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))), size = 3.5) +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11)
  ) +
  scale_fill_gradient(low = "gray95", high = "gray45") +
  labs(
    x = "",
    y = "",
    title = "Quarterly Presentation Type Breakdown",
    subtitle = "Starting with the first complete quarter of presentation-type detail"
  )
```

# Total Dissemination Outputs

```{r dissemination}
dissem <- bind_rows(
  `Scientific publications` = pubsumm,
  `Print/media` = mediasumm,
  Presentations = pressumm,
  .id = "Category"
)

dissem %>%
  group_by(Year, Category) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = Year, y = Quantity, fill = Category)) +
  geom_col(alpha = 0.75) +
  geom_text(
    aes(y = Quantity - 1, label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    position = "stack",
    color = "white",
    size = 3
  ) +
  theme_custom() +
  theme(legend.spacing.x = unit(1, "mm"), legend.box.margin = margin(-20, 0, 0, 0)) +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  labs(x = "", y = "", title = "Number of Dissemination Outputs")
```

# Capacity Sharing & Training

```{r capacity-sharing}
engagement_unique_ind <- metric_summary(
  metrics,
  category = "Capacity Sharing & Training",
  value = "quantity",
  pre2019_category = "Citizen/stakeholder engagement"
)

plot_quarterly_bars(engagement_unique_ind, "Number of People Engaged Through Capacity Sharing & Training", -10)
plot_annual_bars(engagement_unique_ind, "Number of People Engaged Through Capacity Sharing & Training", -10)
```

# Field Research

```{r field-research}
field_days <- metrics %>%
  filter(Category == "Field research") %>%
  group_by(Year, Quarter) %>%
  summarise(Quantity = sum(FieldDays, na.rm = TRUE), .groups = "drop") %>%
  bind_rows(filter(pre2019, Category == "Field research") %>% select(Year, Quarter, Quantity)) %>%
  mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

field_person_days <- metrics %>%
  filter(Category == "Field research") %>%
  group_by(Year, Quarter) %>%
  summarise(Quantity = sum(FieldPersonDays, na.rm = TRUE), .groups = "drop") %>%
  bind_rows(filter(pre2019, Category == "Field research") %>% select(Year, Quarter, Quantity)) %>%
  mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

field_by_department <- metrics %>%
  filter(Category == "Field research") %>%
  select(
    Year,
    Quarter,
    `Conservation Research` = FieldConservationResearch,
    `Animal Health` = FieldAnimalHealth,
    `Animal Care` = FieldAnimalCare,
    `Other Shedd departments` = FieldOtherDepartments
  ) %>%
  pivot_longer(
    cols = -c(Year, Quarter),
    names_to = "Department",
    values_to = "PersonDays"
  ) %>%
  group_by(Year, Quarter, Department) %>%
  summarise(PersonDays = sum(PersonDays, na.rm = TRUE), .groups = "drop")

plot_quarterly_bars(field_days, "Number of Field Days", -5)
plot_annual_bars(field_days, "Number of Field Days", -5)

plot_quarterly_bars(field_person_days, "Number of Field Person-Days", -5)
plot_annual_bars(field_person_days, "Number of Field Person-Days", -5)

field_by_department %>%
  group_by(Year, Department) %>%
  summarise(PersonDays = sum(PersonDays, na.rm = TRUE), .groups = "drop") %>%
  filter(PersonDays > 0) %>%
  ggplot(aes(x = Year, y = PersonDays, fill = Department)) +
  geom_col(alpha = 0.75) +
  theme_custom() +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  labs(x = "", y = "", title = "Field Person-Days by Shedd Department")
```

# Data Quality Checks

```{r data-quality}
metrics_all %>%
  filter(is.na(Category) | is.na(Date)) %>%
  select(Source, ID, Date, RawCategory, Category, Description) %>%
  knitr::kable(caption = "Rows needing category or date review")

metrics %>%
  count(Source, RawCategory, Category, sort = TRUE) %>%
  knitr::kable(caption = "Raw-to-standard category crosswalk")
```
