---
title: "Conservation Research Metrics Report"
subtitle: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 1
date: 
params:
  cutoff: !r Sys.Date()
  refresh_citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 4,
  fig.width = 8
)

cutoff <- as.Date(params$cutoff)
refresh_citations <- isTRUE(params$refresh_citations)
```

```{r report-css, results='asis'}
logo_path <- c(
  "assets/shedd-logo.png",
  "assets/shedd-logo.jpg",
  "assets/shedd-logo.svg",
  "data/shedd-logo.png",
  "data/shedd-logo.jpg",
  "data/shedd-logo.svg"
)
logo_path <- logo_path[file.exists(logo_path)][1]

cat("
<style>
.date {
  font-size: 1.3em;
  font-weight: 500;
}

body {
  font-family: Aptos, Arial, sans-serif;
}


</style>
")

if (!is.na(logo_path)) {
  logo_uri <- knitr::image_uri(logo_path)
  cat(sprintf("
<style>
h1.title::before {
  content: '';
  display: block;
  width: 290px;
  height: 103px;
  margin: 0 0 0.4em 0;
  background-image: url(\"%s\");
  background-position: left center;
  background-repeat: no-repeat;
  background-size: contain;
}
</style>
",
    logo_uri
  ))
}
```

```{r libraries}
library(readxl)
library(tidyverse)
library(lubridate)
library(janitor)
library(ggthemes)
library(RColorBrewer)
library(stringdist)

report_font <- if (
  requireNamespace("systemfonts", quietly = TRUE) &&
    any(systemfonts::system_fonts()$family == "Aptos")
) {
  "Aptos"
} else {
  "Arial"
}

theme_custom <- function() {
  theme_few(base_family = report_font) +
    theme(
      text = element_text(family = report_font),
      legend.title = element_blank(),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.justification = "center",
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 11),
      strip.text = element_text(size = 13),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
    )
}

quarter_levels <- paste0("Q", 1:4)

clean_number <- function(x) {
  as.numeric(str_extract(as.character(x), "-?[0-9]+\\.?[0-9]*"))
}

first_non_missing_date <- function(...) {
  dates <- list(...)
  out <- as.Date(rep(NA, length(dates[[1]])))
  for (date_col in dates) {
    date_col <- as.Date(date_col)
    out <- coalesce(out, date_col)
  }
  out
}

standardize_category <- function(category) {
  category_norm <- str_squish(str_replace_all(category, "\u00a0", " "))

  case_when(
    str_detect(category_norm, regex("citizen/stakeholder|community engagement|community / stakeholder engagement|capacity sharing|capacity", ignore_case = TRUE)) ~ "Capacity Sharing & Training",
    str_detect(category_norm, regex("media", ignore_case = TRUE)) ~ "Media opportunity",
    str_detect(category_norm, regex("field research|field days|field", ignore_case = TRUE)) ~ "Field research",
    str_detect(category_norm, regex("publication|publications", ignore_case = TRUE)) ~ "Publication",
    str_detect(category_norm, regex("presentation|academic|conference|active meeting|meeting|tour", ignore_case = TRUE)) ~ "Presentation",
    TRUE ~ category_norm
  )
}

standardize_presentation_type <- function(type) {
  type_norm <- str_squish(type)

  case_when(
    type_norm %in% c("Panel discussion") ~ "Panel",
    TRUE ~ type_norm
  )
}

metric_summary <- function(data, category, value = c("n", "quantity"), pre2019_category = category) {
  value <- match.arg(value)

  current <- data %>%
    filter(Category == category) %>%
    group_by(Year, Quarter) %>%
    summarise(
      Quantity = if (value == "n") n() else sum(Quantity, na.rm = TRUE),
      .groups = "drop"
    )

  historical <- pre2019 %>%
    filter(Category == pre2019_category) %>%
    transmute(Year, Quarter, Quantity = clean_number(Quantity))

  bind_rows(current, historical) %>%
    mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
    complete(
      Year = full_seq(range(Year, na.rm = TRUE), 1),
      Quarter = factor(quarter_levels, levels = quarter_levels),
      fill = list(Quantity = 0)
    ) %>%
    arrange(Year, Quarter)
}

plot_quarterly_bars <- function(summary_data, title, label_nudge) {
  ggplot(summary_data, aes(x = Quarter, y = Quantity, fill = factor(Year))) +
    facet_grid(~ Year) +
    geom_col(alpha = 0.75) +
    geom_text(
      aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
      color = "white",
      nudge_y = label_nudge,
      size = 3
    ) +
    theme_custom() +
    theme(legend.position = "none", axis.text.x = element_text(size = 7)) +
    labs(x = "", y = "", title = title)
}

plot_annual_bars <- function(summary_data, title, label_nudge) {
  year_breaks <- seq(min(summary_data$Year, na.rm = TRUE), max(summary_data$Year, na.rm = TRUE), by = 1)

  summary_data %>%
    group_by(Year) %>%
    summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop") %>%
    ggplot(aes(x = Year, y = Quantity, fill = factor(Year))) +
    geom_col(alpha = 0.75) +
    geom_text(
      aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
      color = "white",
      nudge_y = label_nudge,
      size = 3
    ) +
    theme_custom() +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = year_breaks) +
    labs(x = "", y = "", title = title)
}

dedupe_references <- function(data, reference_col = Citation, threshold = 0.10) {
  data <- data %>%
    filter(!is.na({{ reference_col }}), {{ reference_col }} != "")

  if (nrow(data) < 2) {
    return(data)
  }

  groups <- data %>%
    pull({{ reference_col }}) %>%
    stringdistmatrix(method = "jw") %>%
    as.dist() %>%
    hclust(method = "average") %>%
    cutree(h = threshold)

  data[!duplicated(groups), ]
}

latest_matching_file <- function(path, pattern) {
  files <- list.files(path = path, pattern = pattern, full.names = TRUE)

  if (length(files) == 0) {
    return(NA_character_)
  }

  files[which.max(file.info(files)$mtime)]
}

read_scival_lines <- function(path) {
  read_lines(path, locale = locale(encoding = "UTF-8")) %>%
    str_remove("^\ufeff") %>%
    str_replace("^\\t+", "")
}

read_scival_metadata <- function(path) {
  metadata_keys <- c(
    "Data set",
    "Entity",
    "Year range",
    "Subject classification",
    "Filtered by",
    "Types of publications included",
    "Self-citations",
    "Data source",
    "Date last updated",
    "Date exported"
  )

  lines <- read_scival_lines(path)

  map_dfr(metadata_keys, function(key) {
    line <- lines[str_starts(lines, fixed(paste0(key, ",")))][1]

    if (is.na(line)) {
      return(tibble(Field = key, Value = NA_character_))
    }

    row <- read.csv(
      text = line,
      header = FALSE,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )

    tibble(Field = key, Value = as.character(row[1, 2]))
  })
}

read_scival_summary_metrics <- function(path) {
  summary_metrics <- c(
    "Scholarly Output cited by Policies",
    "Scholarly Output cited by Policies (%)",
    "Citing Policy Documents",
    "Policy Citations",
    "Policy Bodies",
    "Policy Body Countries",
    "Field-Normalized Policy Impact"
  )

  lines <- read_scival_lines(path)

  map_dfr(summary_metrics, function(metric) {
    line <- lines[str_starts(lines, fixed(paste0(metric, ",")))][1]

    if (is.na(line)) {
      return(tibble(Metric = metric, Value = NA_real_))
    }

    row <- read.csv(
      text = line,
      header = FALSE,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )

    tibble(Metric = metric, Value = as.numeric(row[1, 2]))
  })
}

read_scival_wide_table <- function(path) {
  lines <- read_scival_lines(path)
  header_index <- which(str_detect(lines, "Overall") & str_detect(lines, "\\b(19|20)[0-9]{2}\\b"))[1]

  if (is.na(header_index)) {
    return(tibble())
  }

  data_lines <- lines[header_index:length(lines)]
  data_lines <- data_lines[
    str_squish(data_lines) != "" &
      !str_detect(data_lines, "Elsevier B\\.V\\.")
  ] %>%
    str_replace("^\\t+", "")

  read_csv(
    I(data_lines),
    show_col_types = FALSE,
    name_repair = "unique",
    na = c("", "NA")
  ) %>%
    select(where(~ !all(is.na(.x))))
}

read_scival_series <- function(path) {
  table <- read_scival_wide_table(path)

  if (nrow(table) == 0) {
    return(tibble())
  }

  label_cols <- names(table)[str_detect(names(table), "^\\.\\.\\.")]
  value_cols <- setdiff(names(table), label_cols)

  table %>%
    filter(if_any(everything(), ~ !is.na(.x))) %>%
    mutate(
      Measure = pmap_chr(across(all_of(label_cols)), function(...) {
        labels <- c(...)
        labels <- labels[!is.na(labels) & str_squish(as.character(labels)) != ""]
        str_squish(paste(labels, collapse = " - "))
      })
    ) %>%
    select(Measure, all_of(value_cols)) %>%
    pivot_longer(-Measure, names_to = "Period", values_to = "Value") %>%
    mutate(
      Value = as.numeric(Value),
      Year = suppressWarnings(as.integer(Period))
    ) %>%
    filter(!is.na(Value))
}

read_scival_policy_documents <- function(path) {
  lines <- read_scival_lines(path)
  header_index <- which(str_detect(lines, "^\"Title\",\"Page Source Url\""))[1]

  if (is.na(header_index)) {
    return(tibble())
  }

  read_csv(
    I(lines[header_index:length(lines)]),
    show_col_types = FALSE,
    name_repair = "unique",
    na = c("", "NA")
  ) %>%
    filter(!is.na(Title), Title != "")
}

spread_policy_labels <- function(y, min_gap = 0.17, lower = -1.12, upper = 1.12) {
  if (length(y) == 0) {
    return(y)
  }

  ordered_index <- order(y, decreasing = TRUE)
  adjusted <- pmin(pmax(y[ordered_index], lower), upper)

  if (length(adjusted) > 1) {
    for (i in 2:length(adjusted)) {
      adjusted[[i]] <- min(adjusted[[i]], adjusted[[i - 1]] - min_gap)
    }
  }

  if (tail(adjusted, 1) < lower) {
    adjusted <- adjusted + (lower - tail(adjusted, 1))
  }

  if (adjusted[[1]] > upper) {
    adjusted <- adjusted - (adjusted[[1]] - upper)
  }

  out <- numeric(length(y))
  out[ordered_index] <- adjusted
  out
}

extract_dois <- function(reference) {
  reference %>%
    str_replace_all(regex("%2f", ignore_case = TRUE), "/") %>%
    str_extract_all(regex("10\\.\\d{4,9}/[^\\s\"<>]+", ignore_case = TRUE)) %>%
    unlist(use.names = FALSE) %>%
    str_replace_all(regex("^https?://(dx\\.)?doi\\.org/", ignore_case = TRUE), "") %>%
    str_remove("[?#].*$") %>%
    str_remove(regex("(\\.(full|abstract|pdf|epdf|xml|html))+$", ignore_case = TRUE)) %>%
    str_remove("[\\]\\)\\.,;:]+$") %>%
    str_to_lower()
}

normalize_publication_text <- function(x) {
  x %>%
    str_to_lower() %>%
    str_replace_all("[^a-z0-9]+", " ") %>%
    str_squish()
}

publication_reference_columns <- c("Citation", "Link", "Publication Link")

publication_reference_text <- function(data) {
  available_cols <- intersect(publication_reference_columns, names(data))

  if (length(available_cols) == 0) {
    return(rep(NA_character_, nrow(data)))
  }

  pmap_chr(data[available_cols], function(...) {
    values <- as.character(c(...))
    values <- str_squish(values)
    values <- values[!is.na(values) & values != "" & str_to_upper(values) != "NA"]

    if (length(values) == 0) {
      return(NA_character_)
    }

    str_squish(paste(values, collapse = " "))
  })
}

publication_references <- function(data, category_col) {
  publications <- data %>%
    filter(str_detect(coalesce(as.character(.data[[category_col]]), ""), regex("Publication", ignore_case = TRUE)))

  publications %>%
    transmute(
      reference = as.character(Citation),
      DOIReference = publication_reference_text(publications)
    )
}

first_crossref_year <- function(item) {
  if (is.null(item$issued) || is.null(item$issued[["date-parts"]])) {
    return(NA_integer_)
  }

  date_parts <- item$issued[["date-parts"]][[1]]

  if (length(date_parts) == 0) {
    return(NA_integer_)
  }

  as.integer(date_parts[[1]])
}

or_missing <- function(x, missing_value) {
  if (is.null(x) || length(x) == 0) {
    return(missing_value)
  }

  x
}

resolve_doi_from_crossref <- function(title, year = NA_integer_) {
  if (
    !requireNamespace("httr", quietly = TRUE) ||
      !requireNamespace("jsonlite", quietly = TRUE)
  ) {
    return(tibble(DOI = NA_character_, MatchTitle = NA_character_, MatchScore = NA_real_, MatchYear = NA_integer_))
  }

  response <- tryCatch(
    httr::GET(
      "https://api.crossref.org/works",
      query = list(
        "query.title" = title,
        rows = 5,
        select = "DOI,title,issued,score"
      ),
      httr::user_agent("Shedd research metrics DOI lookup"),
      httr::timeout(15)
    ),
    error = function(e) NULL
  )

  if (is.null(response) || httr::status_code(response) != 200) {
    return(tibble(DOI = NA_character_, MatchTitle = NA_character_, MatchScore = NA_real_, MatchYear = NA_integer_))
  }

  content <- httr::content(response, as = "text", encoding = "UTF-8")
  items <- jsonlite::fromJSON(content, simplifyVector = FALSE)$message$items

  if (length(items) == 0) {
    return(tibble(DOI = NA_character_, MatchTitle = NA_character_, MatchScore = NA_real_, MatchYear = NA_integer_))
  }

  title_norm <- normalize_publication_text(title)

  candidates <- map_dfr(items, function(item) {
    match_title <- if (length(item$title) > 0) item$title[[1]] else NA_character_
    match_year <- first_crossref_year(item)

    tibble(
      DOI = str_to_lower(or_missing(item$DOI, NA_character_)),
      MatchTitle = match_title,
      MatchScore = stringdist::stringsim(title_norm, normalize_publication_text(match_title), method = "jw"),
      MatchYear = match_year,
      CrossrefScore = or_missing(item$score, NA_real_)
    )
  }) %>%
    mutate(YearMatches = !is.na(year) & !is.na(MatchYear) & year == MatchYear) %>%
    filter(!is.na(DOI), MatchScore >= 0.88 | (YearMatches & MatchScore >= 0.82)) %>%
    arrange(desc(YearMatches), desc(MatchScore), desc(CrossrefScore))

  if (nrow(candidates) == 0) {
    return(tibble(DOI = NA_character_, MatchTitle = NA_character_, MatchScore = NA_real_, MatchYear = NA_integer_))
  }

  candidates %>%
    slice(1) %>%
    select(DOI, MatchTitle, MatchScore, MatchYear)
}

doi_from_references <- function(title, reference_lookup) {
  title_norm <- normalize_publication_text(title)
  matches <- reference_lookup %>%
    filter(str_detect(ReferenceNorm, fixed(title_norm))) %>%
    pull(DOI) %>%
    unlist(use.names = FALSE)

  matches <- matches[!is.na(matches) & matches != ""]

  if (length(matches) == 0) {
    return(NA_character_)
  }

  matches[1]
}

write_publication_dois <- function(
  publications,
  references,
  output_file = "output/publication_dois.txt",
  detail_file = "output/publication_doi_details.csv"
) {
  reference_lookup <- references

  if (!"DOIReference" %in% names(reference_lookup)) {
    reference_lookup <- reference_lookup %>%
      mutate(DOIReference = reference)
  }

  reference_lookup <- reference_lookup %>%
    mutate(
      ReferenceNorm = normalize_publication_text(reference),
      DOIReference = str_squish(as.character(DOIReference)),
      DOIReference = if_else(str_to_upper(DOIReference) == "NA", NA_character_, DOIReference),
      DOIReference = coalesce(na_if(DOIReference, ""), as.character(reference)),
      DOI = map(DOIReference, extract_dois)
    )

  publication_dois <- publications %>%
    filter(!is.na(title), title != "") %>%
    mutate(
      ScholarAuthorID = if ("scholar_id" %in% names(.)) as.character(scholar_id) else NA_character_,
      ScholarCID = if ("cid" %in% names(.)) as.character(cid) else NA_character_,
      ScholarPubID = if ("pubid" %in% names(.)) as.character(pubid) else NA_character_
    ) %>%
    transmute(
      Title = title,
      Year = as.integer(year),
      ScholarAuthorID,
      ScholarCID,
      ScholarPubID,
      DOI = map_chr(title, doi_from_references, reference_lookup),
      DOISource = if_else(is.na(DOI), NA_character_, "local reference")
    )

  if (file.exists(detail_file)) {
    cached_dois <- read_csv(detail_file, show_col_types = FALSE) %>%
      filter(!is.na(DOI), DOI != "") %>%
      transmute(
        TitleNorm = normalize_publication_text(Title),
        Year = as.integer(Year),
        CachedDOI = DOI,
        CachedDOISource = DOISource
      ) %>%
      distinct(TitleNorm, Year, .keep_all = TRUE)

    publication_dois <- publication_dois %>%
      mutate(TitleNorm = normalize_publication_text(Title)) %>%
      left_join(cached_dois, by = c("TitleNorm", "Year")) %>%
      mutate(
        DOI = coalesce(DOI, CachedDOI),
        DOISource = coalesce(DOISource, CachedDOISource)
      ) %>%
      select(-TitleNorm, -CachedDOI, -CachedDOISource)
  }

  missing_doi_rows <- which(is.na(publication_dois$DOI))

  if (length(missing_doi_rows) > 0) {
    crossref_matches <- map_dfr(missing_doi_rows, function(row_index) {
      Sys.sleep(0.1)

      resolve_doi_from_crossref(
        publication_dois$Title[[row_index]],
        publication_dois$Year[[row_index]]
      ) %>%
        mutate(RowIndex = row_index)
    })

    if (nrow(crossref_matches) > 0) {
      for (match_index in seq_len(nrow(crossref_matches))) {
        row_index <- crossref_matches$RowIndex[[match_index]]
        doi <- crossref_matches$DOI[[match_index]]

        if (!is.na(doi) && doi != "") {
          publication_dois$DOI[[row_index]] <- doi
          publication_dois$DOISource[[row_index]] <- "Crossref title lookup"
        }
      }
    }
  }

  doi_list <- publication_dois %>%
    filter(!is.na(DOI), DOI != "") %>%
    distinct(DOI) %>%
    arrange(DOI) %>%
    pull(DOI)

  dir.create(dirname(output_file), showWarnings = FALSE, recursive = TRUE)
  write_lines(doi_list, output_file)
  write_csv(publication_dois, detail_file)

  tibble(
    DOIFile = output_file,
    DetailFile = detail_file,
    Publications = nrow(publication_dois),
    UniqueDOIs = length(doi_list)
  )
}
```

<!-- # Import And Normalize Data -->

```{r import-data, include=FALSE}
pre2019 <- read_xlsx("data/pre2019_summary.xlsx") %>%
  mutate(
    Year = as.integer(Year),
    Quarter = factor(Quarter, levels = quarter_levels),
    Quantity = clean_number(Quantity)
  )

metrics_2019_2022 <- read_xlsx("data/research_metrics_2019-2022.xlsx") %>%
  transmute(
    Source = "2019-2022 Microsoft Form",
    ID = as.character(ID),
    Email,
    Name,
    Department,
    Date = as.Date(`Date of event`),
    RawCategory = Category,
    Category = standardize_category(Category),
    MediaOpportunityType = NA_character_,
    PresentationType = NA_character_,
    CapacitySharingType = NA_character_,
    Citation,
    Link,
    Effort = clean_number(Effort),
    Quantity = clean_number(Quantity),
    FieldDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldPersonDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldConservationResearch = if_else(Category == "Field research", Quantity, NA_real_),
    FieldAnimalHealth = NA_real_,
    FieldAnimalCare = NA_real_,
    FieldOtherDepartments = NA_real_,
    Description,
    Program,
    Pillar
  )

metrics_2022_2024_raw <- read_xlsx(
  "data/research_metrics_2022-2024.xlsx",
  col_types = c(rep("guess", 8), "text", "text", rep("guess", 26))
)

metrics_2022_2024 <- metrics_2022_2024_raw %>%
  transmute(
    Source = "2022-2024 IT Form",
    ID = as.character(ID),
    Email,
    Name,
    Department = `Select your Animal Care & Science department`,
    SingleMulti = `Was this a single-day or multi-day engagement?`,
    DateEngagement = as.Date(`What was the date of the engagement?`),
    DateStart = as.Date(`What was the starting date of the engagement?`),
    DateEnd = as.Date(`What was the ending date of the engagement?`),
    DatePublication = as.Date(`Publication Date`),
    DateMedia = as.Date(`Media Opportunity Date`),
    RawCategory = `What was the engagement type?`,
    Category = standardize_category(RawCategory),
    MediaOpportunityType = NA_character_,
    PresentationType = NA_character_,
    CapacitySharingType = NA_character_,
    Citation,
    Link = case_when(
      Category == "Publication" ~ `Publication Link`,
      Category == "Media opportunity" ~ `Media Link`,
      TRUE ~ NA_character_
    ),
    Effort = clean_number(`Community / Stakeholder Engagement Effort`),
    Quantity = case_when(
      Category == "Field research" & SingleMulti == "Multi-day" ~ as.numeric(DateEnd - DateStart) + 1,
      Category == "Field research" & SingleMulti == "Single-day" ~ 1,
      Category == "Capacity Sharing & Training" ~ clean_number(`Community / Stakeholder Engagement Quantity`),
      TRUE ~ NA_real_
    ),
    FieldDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldPersonDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldConservationResearch = if_else(Category == "Field research", Quantity, NA_real_),
    FieldAnimalHealth = NA_real_,
    FieldAnimalCare = NA_real_,
    FieldOtherDepartments = NA_real_,
    Description = `Briefly describe, or enter any notes for, this engagement`,
    Program = `If the engagement is for a program, select it from the list below`,
    Pillar = NA_character_,
    Date = first_non_missing_date(DatePublication, DateMedia, DateEngagement, DateStart, as.Date(`Start time`))
  ) %>%
  select(-SingleMulti, -DateEngagement, -DateStart, -DateEnd, -DatePublication, -DateMedia)

onedrive_metrics_file <- c(
  file.path(
    Sys.getenv("HOME"),
    "Library",
    "CloudStorage",
    "OneDrive-SheddAquariumSociety",
    "Research metrics new 1.xlsx"
  ),
  file.path(
    Sys.getenv("HOME"),
    "OneDrive - Shedd Aquarium Society",
    "Research metrics new 1.xlsx"
  )
)
onedrive_metrics_file <- onedrive_metrics_file[file.exists(onedrive_metrics_file)][1]

if (!is.na(onedrive_metrics_file)) {
  local_metrics_file <- file.path("data", basename(onedrive_metrics_file))
  onedrive_metrics_info <- file.info(onedrive_metrics_file)
  onedrive_metrics_available <- !is.na(onedrive_metrics_info$size) && onedrive_metrics_info$size > 0

  if (
    onedrive_metrics_available &&
      (
        !file.exists(local_metrics_file) ||
          file.info(local_metrics_file)$size == 0 ||
          onedrive_metrics_info$mtime > file.info(local_metrics_file)$mtime
      )
  ) {
    temp_metrics_file <- tempfile(fileext = ".xlsx")
    copied_metrics_file <- file.copy(onedrive_metrics_file, temp_metrics_file, overwrite = TRUE)
    copied_metrics_available <- copied_metrics_file &&
      !is.na(file.info(temp_metrics_file)$size) &&
      file.info(temp_metrics_file)$size > 0

    if (copied_metrics_available) {
      file.copy(temp_metrics_file, local_metrics_file, overwrite = TRUE)
    } else {
      warning("Unable to copy latest OneDrive metrics workbook; using the local data/ copy.")
    }

    unlink(temp_metrics_file)
  }
}

new_metrics_candidates <- list.files(path = "data", pattern = "^Research metrics new.*\\.xlsx$", full.names = TRUE)
new_metrics_candidates <- new_metrics_candidates[file.info(new_metrics_candidates)$size > 0]

if (length(new_metrics_candidates) == 0) {
  stop("No non-empty 'Research metrics new*.xlsx' file found in data/ or the configured OneDrive root.")
}

new_metrics_file <- new_metrics_candidates[which.max(file.info(new_metrics_candidates)$mtime)]

metrics_2024_on <- read_xlsx(new_metrics_file) %>%
  transmute(
    Source = "2024-on Microsoft Form",
    ID = as.character(Id),
    Email,
    Name,
    Department,
    Date = as.Date(`Date of event`),
    RawCategory = Category,
    Category = standardize_category(Category),
    MediaOpportunityType = str_squish(`Type of media opportunity`),
    PresentationType = standardize_presentation_type(`Type of presentation`),
    CapacitySharingType = str_squish(`Type of capacity sharing & training`),
    Citation,
    Link,
    Effort = clean_number(coalesce(as.character(Effort), as.character(Effort1))),
    Quantity = clean_number(Quantity),
    FieldDays = if_else(Category == "Field research", Quantity, NA_real_),
    FieldConservationResearchHeadcount = clean_number(`Field Team: Number of members from Conservation Research`),
    FieldAnimalHealthHeadcount = clean_number(`Field Team: Number of members from Animal Health`),
    FieldAnimalCareHeadcount = clean_number(`Field Team: Number of members from Animal Care`),
    FieldOtherDepartmentsHeadcount = clean_number(`Field Team: Number of members from other Shedd Departments`),
    FieldTeamHeadcount = rowSums(
      across(ends_with("Headcount"), ~ replace_na(.x, 0)),
      na.rm = TRUE
    ),
    FieldConservationResearch = case_when(
      Category == "Field research" & FieldTeamHeadcount > 0 ~ Quantity * FieldConservationResearchHeadcount,
      Category == "Field research" ~ Quantity,
      TRUE ~ NA_real_
    ),
    FieldAnimalHealth = if_else(Category == "Field research", Quantity * replace_na(FieldAnimalHealthHeadcount, 0), NA_real_),
    FieldAnimalCare = if_else(Category == "Field research", Quantity * replace_na(FieldAnimalCareHeadcount, 0), NA_real_),
    FieldOtherDepartments = if_else(Category == "Field research", Quantity * replace_na(FieldOtherDepartmentsHeadcount, 0), NA_real_),
    FieldPersonDays = case_when(
      Category == "Field research" & FieldTeamHeadcount > 0 ~ Quantity * FieldTeamHeadcount,
      Category == "Field research" ~ Quantity,
      TRUE ~ NA_real_
    ),
    Description,
    Program,
    Pillar
  ) %>%
  select(-ends_with("Headcount"), -FieldTeamHeadcount)

metrics_combined_with_keys <- bind_rows(metrics_2019_2022, metrics_2022_2024, metrics_2024_on) %>%
  mutate(
    Year = year(Date),
    Quarter = factor(paste0("Q", quarter(Date)), levels = quarter_levels),
    Department = coalesce(Department, "Conservation Research")
  ) %>%
  mutate(
    DescriptionKey = str_squish(str_to_lower(coalesce(Description, ""))),
    CitationKey = str_squish(str_to_lower(coalesce(Citation, ""))),
    LinkKey = str_squish(str_to_lower(coalesce(Link, ""))),
    DuplicateKey = case_when(
      Category == "Publication" & CitationKey != "" ~ paste(Category, CitationKey),
      LinkKey != "" ~ paste(Category, Date, LinkKey),
      DescriptionKey != "" ~ paste(Category, Date, DescriptionKey),
      TRUE ~ paste(Category, Date, Name, Quantity)
    )
  )

metrics_duplicates_removed <- metrics_combined_with_keys %>%
  filter(duplicated(DuplicateKey))

metrics_all <- metrics_combined_with_keys %>%
  filter(!duplicated(DuplicateKey)) %>%
  select(
    -DuplicateKey,
    -DescriptionKey,
    -CitationKey,
    -LinkKey
  )

metrics <- metrics_all %>%
  filter(!is.na(Date), Date <= cutoff)
```

```{r import-policy-data, include=FALSE}
policy_dir <- "data/policy"
policy_documents_file <- file.path(policy_dir, "Policies.csv")

policy_files <- c(
  summary = latest_matching_file(policy_dir, "^Summary_Metrics.*\\.csv$"),
  citing_documents = latest_matching_file(policy_dir, "^Citing_Policy_Documents.*\\.csv$"),
  policies = if (file.exists(policy_documents_file)) policy_documents_file else NA_character_
)

policy_enabled <- dir.exists(policy_dir) && !any(is.na(policy_files))

if (policy_enabled) {
  policy_metadata <- read_scival_metadata(policy_files[["summary"]]) %>%
    mutate(Export = "summary", File = basename(policy_files[["summary"]]), .before = Field)

  policy_summary_metrics <- read_scival_summary_metrics(policy_files[["summary"]])
  policy_document_counts_series <- read_scival_series(policy_files[["citing_documents"]])
} else {
  policy_metadata <- tibble(Export = character(), File = character(), Field = character(), Value = character())
  policy_summary_metrics <- tibble(Metric = character(), Value = numeric())
  policy_document_counts_series <- tibble(Measure = character(), Period = character(), Value = numeric(), Year = integer())
}

policy_citing_documents <- if (!is.na(policy_files[["policies"]]) && file.exists(policy_files[["policies"]])) {
  read_scival_policy_documents(policy_files[["policies"]])
} else {
  tibble()
}
```

```{r import-audit, include=FALSE}
current_source_audit <- metrics %>%
  group_by(Source) %>%
  summarise(
    StartDate = min(Date, na.rm = TRUE),
    EndDate = max(Date, na.rm = TRUE),
    Entries = n(),
    .groups = "drop"
  ) %>%
  mutate(
    DataFile = recode(
      Source,
      `2019-2022 Microsoft Form` = "data/research_metrics_2019-2022.xlsx",
      `2022-2024 IT Form` = "data/research_metrics_2022-2024.xlsx",
      `2024-on Microsoft Form` = new_metrics_file
    ),
    Maintainer = recode(
      Source,
      `2019-2022 Microsoft Form` = "Conservation Research (R. Cunning)",
      `2022-2024 IT Form` = "IT (A. Ingram)",
      `2024-on Microsoft Form` = "Conservation Research (R. Cunning)"
    )
  )

historical_source_audit <- pre2019 %>%
  summarise(
    Source = "Pre-2019 historical summary",
    StartDate = as.Date(paste0(min(Year, na.rm = TRUE), "-01-01")),
    EndDate = as.Date(paste0(max(Year, na.rm = TRUE), "-12-31")),
    Entries = n_distinct(paste(Year, Quarter, Category)),
    DataFile = "data/pre2019_summary.xlsx",
    Maintainer = "Historical summary workbook"
  )

policy_source_audit <- if (policy_enabled) {
  policy_year_ranges <- policy_metadata %>%
    filter(Field == "Year range", !is.na(Value)) %>%
    distinct(Value) %>%
    pull(Value) %>%
    paste(collapse = "; ")

  tibble(
    DataFile = "data/policy/Summary_Metrics*.csv; data/policy/Citing_Policy_Documents*.csv; data/policy/Policies.csv",
    DateRange = policy_year_ranges,
    Maintainer = "SciVal export (Shedd Conservation Research entity)",
    Entries = sum(!is.na(policy_files))
  )
} else {
  tibble(DataFile = character(), DateRange = character(), Maintainer = character(), Entries = integer())
}

bind_rows(
  current_source_audit %>%
    mutate(DateRange = paste(format(StartDate, "%Y-%m-%d"), format(EndDate, "%Y-%m-%d"), sep = " to ")) %>%
    select(DataFile, DateRange, Maintainer, Entries),
  historical_source_audit %>%
    mutate(DateRange = paste(format(StartDate, "%Y-%m-%d"), format(EndDate, "%Y-%m-%d"), sep = " to ")) %>%
    select(DataFile, DateRange, Maintainer, Entries),
  policy_source_audit
) %>%
  transmute(
    `Data file` = DataFile,
    `Date range` = DateRange,
    Maintainer,
    `Entries retained` = Entries
  ) %>%
  arrange(`Date range`) %>%
  knitr::kable(caption = "Metric data sources used in this report")
```

<!--
## Working Assumptions

This overhaul standardizes historical metric names to the current reporting
names. `Citizen/stakeholder engagement`, `Community / Stakeholder Engagement`,
`Community Engagement`, and `Capacity Sharing & Training` are all reported as
`Capacity Sharing & Training`, with the reported quantity interpreted as number
of people engaged. `Conference or active meeting`, `Conference/Meeting`,
`Academic`, `Tour`, and other presentation-like labels are reported as
`Presentation`. `Media` is reported as `Media opportunity`, and old/new field
labels are reported as `Field research`.

For field research, older rows without team headcounts are carried forward as
both field days and Conservation Research person-days. Newer rows with
department headcounts use:

`field person-days = field days * total Shedd field team headcount`
-->

---

# Metrics at a Glance

```{r citation-data}
citation_cache <- "data/citation_counts.txt"
scholar_publication_cache <- "output/google_scholar_publications.csv"

citation_history <- read.delim(
  citation_cache,
  header = TRUE,
  sep = "\t",
  fill = TRUE,
  stringsAsFactors = FALSE
) %>%
  mutate(CacheRow = row_number()) %>%
  as_tibble() %>%
  clean_names() %>%
  mutate(date = as.Date(date))

hc_oldrefs <- read_tsv("data/pre2019_publications.txt", col_names = "reference", show_col_types = FALSE) %>%
  mutate(DOIReference = reference)
hc_newrefs <- metrics %>%
  publication_references("Category")

hc_references <- bind_rows(hc_oldrefs, hc_newrefs) %>%
  dedupe_references(reference)

needs_scholar_pull <- refresh_citations &&
  (
    !Sys.Date() %in% citation_history$date ||
      !file.exists(scholar_publication_cache)
  )

if (needs_scholar_pull) {
  library(scholar)

  hc_authors <- read_tsv("data/authors.txt", show_col_types = FALSE) %>% drop_na()

  all_pubs <- hc_authors %>%
    mutate(all_pubs = map(scholar_id, get_publications)) %>%
    unnest(all_pubs)

  hc_pubs <- all_pubs %>%
    mutate(hc_affil = map_lgl(title, ~ any(agrepl(.x, hc_references$reference)))) %>%
    filter(hc_affil, !is.na(year)) %>%
    filter(!(duplicated(cid) & duplicated(tolower(title))))

  dir.create(dirname(scholar_publication_cache), showWarnings = FALSE, recursive = TRUE)
  write_csv(hc_pubs, scholar_publication_cache)
} else if (file.exists(scholar_publication_cache)) {
  hc_pubs <- read_csv(
    scholar_publication_cache,
    col_types = cols(
      .default = col_guess(),
      cid = col_character(),
      pubid = col_character()
    )
  )
} else {
  hc_pubs <- tibble()
}

if (nrow(hc_pubs) > 0) {
  publication_doi_summary <- write_publication_dois(hc_pubs, hc_references)
} else {
  publication_doi_summary <- tibble(
    DOIFile = "output/publication_dois.txt",
    DetailFile = "output/publication_doi_details.csv",
    Publications = 0,
    UniqueDOIs = 0
  )
}

if (refresh_citations && !Sys.Date() %in% citation_history$date) {
  pub_cites <- hc_pubs %>%
    arrange(desc(cites)) %>%
    pull(cites)

  h.index <- tail(which(pub_cites >= seq_along(pub_cites)), 1)
  i10.index <- sum(pub_cites >= 10)

  todays_totals <- hc_pubs %>%
    summarise(
      date = Sys.Date(),
      total_pubs = nrow(hc_pubs),
      total_cites = sum(cites),
      h_index = h.index,
      i10_index = i10.index
    )

  bind_rows(citation_history, todays_totals) %>%
    select(-cache_row) %>%
    arrange(date) %>%
    write_tsv(citation_cache)

  citation_history <- read.delim(
    citation_cache,
    header = TRUE,
    sep = "\t",
    fill = TRUE,
    stringsAsFactors = FALSE
  ) %>%
    mutate(CacheRow = row_number()) %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(date = as.Date(date))
}

cites_raw <- citation_history %>%
  group_by(date) %>%
  slice_max(cache_row, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-cache_row) %>%
  filter(date <= cutoff)

latest <- cites_raw %>%
  arrange(date) %>%
  tail(1)

current_year <- year(cutoff)
current_quarter <- factor(paste0("Q", quarter(cutoff)), levels = quarter_levels)

metric_colors <- c(
  publications = "#2C7FB8",
  citations = "#2CA25F",
  policy = "#6A7F2A",
  media = "#F39C12",
  presentations = "#8E44AD",
  capacity = "#D95F02",
  field = "#008B8B"
)

lighten_color <- function(color, amount = 0.55) {
  rgb <- grDevices::col2rgb(color) / 255
  rgb <- rgb * (1 - amount) + amount
  grDevices::rgb(rgb[1], rgb[2], rgb[3])
}

period_fill <- function(metric) {
  c(
    "Complete period" = metric_colors[[metric]],
    "Year to date" = lighten_color(metric_colors[[metric]], 0.45),
    "Quarter to date" = lighten_color(metric_colors[[metric]], 0.45)
  )
}
```

*Report cutoff: `r cutoff`*  
*Latest cached citation pull: `r latest$date`*


```{r metrics-at-a-glance}
publication_start_year <- read_lines("data/pre2019_publications.txt") %>%
  str_extract("\\b20[0-9]{2}\\b") %>%
  as.integer() %>%
  min(na.rm = TRUE)

activity_start_year <- min(pre2019$Year, metrics$Year, na.rm = TRUE)

current_publications <- metrics %>%
  filter(Category == "Publication", Year == current_year) %>%
  dedupe_references(Citation) %>%
  nrow()

media_summary <- metric_summary(metrics, category = "Media opportunity", value = "n")
presentation_summary <- metric_summary(metrics, category = "Presentation", value = "n")
capacity_summary <- metric_summary(
  metrics,
  category = "Capacity Sharing & Training",
  value = "quantity",
  pre2019_category = "Citizen/stakeholder engagement"
)
field_person_days_summary <- metrics %>%
  filter(Category == "Field research") %>%
  group_by(Year, Quarter) %>%
  summarise(Quantity = sum(FieldPersonDays, na.rm = TRUE), .groups = "drop") %>%
  bind_rows(
    filter(pre2019, Category == "Field research") %>%
      transmute(Year, Quarter, Quantity = clean_number(Quantity))
  ) %>%
  mutate(Quarter = factor(Quarter, levels = quarter_levels))

citation_baseline <- cites_raw %>%
  filter(date < as.Date(paste0(current_year, "-01-01"))) %>%
  arrange(date) %>%
  tail(1)

citation_ytd <- if (nrow(citation_baseline) == 0) {
  NA_real_
} else {
  latest$total_cites - citation_baseline$total_cites
}

snapshot_metrics <- tibble(
  MetricKey = c(
    "publications",
    "citations",
    "media",
    "presentations",
    "capacity",
    "field"
  ),
  Metric = c(
    "Scientific publications",
    "Citations",
    "Media opportunities",
    "Presentations",
    "People engaged through capacity sharing & training",
    "Field research person-days"
  ),
  `Current YTD` = c(
    scales::comma(current_publications),
    if_else(is.na(citation_ytd), NA_character_, scales::comma(citation_ytd)),
    scales::comma(sum(media_summary$Quantity[media_summary$Year == current_year], na.rm = TRUE)),
    scales::comma(sum(presentation_summary$Quantity[presentation_summary$Year == current_year], na.rm = TRUE)),
    scales::comma(sum(capacity_summary$Quantity[capacity_summary$Year == current_year], na.rm = TRUE)),
    scales::comma(round(sum(field_person_days_summary$Quantity[field_person_days_summary$Year == current_year], na.rm = TRUE), 0))
  ),
  `Overall total` = c(
    scales::comma(latest$total_pubs),
    scales::comma(latest$total_cites),
    scales::comma(sum(media_summary$Quantity, na.rm = TRUE)),
    scales::comma(sum(presentation_summary$Quantity, na.rm = TRUE)),
    scales::comma(sum(capacity_summary$Quantity, na.rm = TRUE)),
    scales::comma(round(sum(field_person_days_summary$Quantity, na.rm = TRUE), 0))
  ),
  `Overall period` = c(
    paste0(publication_start_year, "-", current_year),
    paste0(publication_start_year, "-", current_year),
    paste0(activity_start_year, "-", current_year),
    paste0(activity_start_year, "-", current_year),
    paste0(activity_start_year, "-", current_year),
    paste0(activity_start_year, "-", current_year)
  )
) %>%
  mutate(
    RowColor = map_chr(MetricKey, ~ lighten_color(metric_colors[[.x]], 0.45))
  )

snapshot_table <- snapshot_metrics %>%
  select(-MetricKey, -RowColor) %>%
  knitr::kable(
    format = "html",
    align = c("l", "r", "r", "l"),
    caption = "Current and cumulative research metrics",
    escape = TRUE
  ) %>%
  kableExtra::kable_styling(
    full_width = TRUE,
    font_size = 16
  )

for (row_index in seq_len(nrow(snapshot_metrics))) {
  snapshot_table <- kableExtra::row_spec(
    snapshot_table,
    row_index,
    color = "black",
    background = snapshot_metrics$RowColor[row_index]
  )
}

snapshot_table
```

---

# Publications & Citations

## Current Publication Impact Snapshot

Latest cached citation pull: `r latest$date`

```{r citation-snapshot}
tibble(
  Metric = c("Publications", "Citations", "h-index", "i10-index"),
  `Current value` = c(
    scales::comma(latest$total_pubs),
    scales::comma(latest$total_cites),
    as.character(latest$h_index),
    as.character(latest$i10_index)
  )
) %>%
  knitr::kable(align = c("l", "r"))
```

## Publications

```{r publications}
nondupes <- metrics %>%
  filter(Category == "Publication") %>%
  dedupe_references(Citation)

pub_historical_annual <- read_lines("data/pre2019_publications.txt") %>%
  tibble(reference = .) %>%
  mutate(Year = as.integer(str_extract(reference, "\\((20[0-9]{2})\\)", group = 1))) %>%
  filter(!is.na(Year), Year <= 2018) %>%
  count(Year, name = "Quantity")

pub_quarterly_current <- nondupes %>%
  filter(Year >= 2019) %>%
  group_by(Year, Quarter) %>%
  summarise(Quantity = n(), .groups = "drop") %>%
  mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

pub_annual <- bind_rows(
  pub_historical_annual,
  pub_quarterly_current %>%
    group_by(Year) %>%
    summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")
) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    fill = list(Quantity = 0)
  )

pubsumm <- bind_rows(
  pub_historical_annual %>%
    mutate(Quarter = factor(NA_character_, levels = quarter_levels)),
  pub_quarterly_current
)

pub_recent_years <- pub_quarterly_current %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

pub_quarterly_recent <- pub_quarterly_current %>%
  filter(Year %in% pub_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

pub_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(pub_annual$Year), max(pub_annual$Year), by = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("publications")[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Number of Scientific Publications")

pub_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("publications")[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly Scientific Publications, Most Recent 5 Years")
```

## Citations

```{r citations}
citesumm <- cites_raw %>%
  mutate(
    Year = year(date),
    Quarter = factor(paste0("Q", quarter(date)), levels = quarter_levels)
  ) %>%
  group_by(Year, Quarter) %>%
  distinct(date, .keep_all = TRUE) %>%
  filter(date == max(date)) %>%
  ungroup() %>%
  arrange(Year, Quarter) %>%
  mutate(Quantity = total_cites - lag(total_cites)) %>%
  transmute(Year, Quarter, Quantity = replace_na(Quantity, 0)) %>%
  bind_rows(
    pre2019 %>%
      filter(Category == "Citations") %>%
      select(Year, Quarter, Quantity)
  ) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

cite_annual <- citesumm %>%
  group_by(Year) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")

cite_recent_years <- citesumm %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

cite_quarterly_recent <- citesumm %>%
  filter(Year %in% cite_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

cite_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(cite_annual$Year), max(cite_annual$Year), by = 1)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("citations")[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Number of New Citations")

cite_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("citations")[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly New Citations, Most Recent 5 Years")
```

---

# Policy Impact

```{r policy-impact-summary, results='asis'}
if (!policy_enabled) {
  cat("_No SciVal Policy Impact exports were found in `data/policy/`._")
} else {
  policy_meta_value <- function(export, field) {
    value <- policy_metadata %>%
      filter(Export == export, Field == field) %>%
      pull(Value)

    if (length(value) == 0 || is.na(value[1])) {
      return("not listed")
    }

    value[1]
  }

  cat(sprintf(
    "*SciVal data last updated: %s*  \n*Exported: %s; summary publication-year range: %s*\n\n",
    policy_meta_value("summary", "Date last updated"),
    policy_meta_value("summary", "Date exported"),
    policy_meta_value("summary", "Year range")
  ))

  policy_summary_display <- policy_summary_metrics %>%
    mutate(
      Metric = recode(
        Metric,
        `Scholarly Output cited by Policies` = "Scholarly output cited by policy documents",
        `Scholarly Output cited by Policies (%)` = "Scholarly output cited by policy documents (%)",
        `Citing Policy Documents` = "Citing policy documents",
        `Policy Citations` = "Policy citations",
        `Policy Bodies` = "Policy bodies",
        `Policy Body Countries` = "Policy body countries",
        `Field-Normalized Policy Impact` = "Field-normalized policy impact"
      ),
      `Current value` = case_when(
        str_detect(Metric, "%|impact") ~ scales::number(Value, accuracy = 0.1),
        TRUE ~ scales::comma(Value, accuracy = 1)
      )
    ) %>%
    select(Metric, `Current value`)

  print(knitr::kable(
    policy_summary_display,
    align = c("l", "r"),
    caption = "SciVal Policy Impact summary for the Shedd Conservation Research entity"
  ))
}
```

```{r policy-impact-citing-documents}
if (policy_enabled) {
  policy_document_counts_annual <- policy_document_counts_series %>%
    filter(Measure == "Citing Policy Documents", !is.na(Year), Year <= current_year) %>%
    transmute(Year, Quantity = Value)

  if (nrow(policy_document_counts_annual) > 0) {
    policy_document_counts_annual <- policy_document_counts_annual %>%
      complete(
        Year = full_seq(range(Year, na.rm = TRUE), 1),
        fill = list(Quantity = 0)
      )

    first_policy_document_year <- min(policy_document_counts_annual$Year[policy_document_counts_annual$Quantity > 0], na.rm = TRUE)

    if (is.finite(first_policy_document_year)) {
      policy_document_counts_annual <- policy_document_counts_annual %>%
        filter(Year >= first_policy_document_year)
    }

    print(
      policy_document_counts_annual %>%
        mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
        ggplot(aes(x = Year, y = Quantity)) +
        geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
        geom_text(
          aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
          color = "black",
          vjust = -0.35,
          size = 3
        ) +
        theme_custom() +
        theme(legend.position = "bottom") +
        scale_x_continuous(
          breaks = seq(
            min(policy_document_counts_annual$Year),
            max(policy_document_counts_annual$Year),
            by = 1
          )
        ) +
        scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
        scale_fill_manual(values = period_fill("citations")[c("Complete period", "Year to date")]) +
        labs(x = "", y = "", title = "Citing Policy Documents by Year")
    )
  }
}
```

```{r policy-impact-location, fig.width=8, fig.height=4}
if (policy_enabled && nrow(policy_citing_documents) > 0) {
    policy_document_locations <- policy_citing_documents %>%
      distinct(
        Title,
        `Page Source Url`,
        `PDF Document Url`,
        Year,
        `Full date`,
        `Policy Body`,
        `Policy Body Country/Region`,
        `Policy Body Type`
      ) %>%
      count(`Policy Body Country/Region`, name = "documents", sort = TRUE) %>%
      mutate(region = if_else(documents > 1, `Policy Body Country/Region`, "Other regions")) %>%
      group_by(region) %>%
      summarise(documents = sum(documents), .groups = "drop") %>%
      arrange(desc(documents)) %>%
      mutate(
        region = factor(region, levels = region),
        fraction = documents / sum(documents),
        percent = scales::percent(fraction, accuracy = 0.1),
        end = cumsum(fraction),
        start = lag(end, default = 0),
        mid = (start + end) / 2,
        angle_start = pi / 2 - 2 * pi * start,
        angle_end = pi / 2 - 2 * pi * end,
        angle_mid = pi / 2 - 2 * pi * mid,
        side = if_else(cos(angle_mid) >= 0, "right", "left"),
        x0 = 1.02 * cos(angle_mid),
        y0 = 1.02 * sin(angle_mid),
        elbow_x = if_else(side == "right", 1.18, -1.18),
        label_x = if_else(side == "right", 1.34, -1.34),
        hjust = if_else(side == "right", 0, 1),
        label = paste0(region, ": ", documents, " (", percent, ")")
      ) %>%
      group_by(side) %>%
      mutate(label_y = spread_policy_labels(y0, min_gap = 0.17, lower = -1.12, upper = 1.12)) %>%
      ungroup()

    policy_location_wedges <- policy_document_locations %>%
      rowwise() %>%
      do({
        theta <- seq(.$angle_start, .$angle_end, length.out = 100)

        tibble(
          region = .$region,
          x = c(0, cos(theta), 0),
          y = c(0, sin(theta), 0)
        )
      }) %>%
      ungroup()

    policy_location_palette <- c(
      "#2C7FB8",
      "#2CA25F",
      "#F39C12",
      "#6A7F2A",
      "#D95F02",
      "#756BB1",
      "#E6550D",
      "#31A354",
      "#969696",
      "#3182BD",
      "#9E9AC8",
      "#FD8D3C",
      "#74C476",
      "#8E44AD",
      "#636363"
    )
    policy_location_palette <- rep(
      policy_location_palette,
      length.out = length(levels(policy_document_locations$region))
    )
    names(policy_location_palette) <- levels(policy_document_locations$region)

    if ("Other regions" %in% names(policy_location_palette)) {
      policy_location_palette[["Other regions"]] <- "#969696"
    }

    print(
      ggplot() +
        geom_polygon(
          data = policy_location_wedges,
          aes(x = x, y = y, group = region, fill = region),
          color = "white",
          linewidth = 0.9
        ) +
        geom_segment(
          data = policy_document_locations,
          aes(x = x0, y = y0, xend = elbow_x, yend = label_y),
          color = "grey50",
          linewidth = 0.3
        ) +
        geom_segment(
          data = policy_document_locations,
          aes(
            x = elbow_x,
            y = label_y,
            xend = label_x - if_else(side == "right", 0.025, -0.025),
            yend = label_y
          ),
          color = "grey50",
          linewidth = 0.3
        ) +
        geom_text(
          data = policy_document_locations,
          aes(x = label_x, y = label_y, label = label, hjust = hjust),
          size = 2.95,
          color = "black"
        ) +
        coord_equal(xlim = c(-3.47, 3.47), ylim = c(-1.28, 1.28), clip = "off") +
        scale_fill_manual(values = policy_location_palette, guide = "none") +
        theme_void(base_size = 12) +
        theme(
          text = element_text(family = report_font),
          panel.border = element_rect(color = "#4D4D4D", fill = NA),
          plot.title = element_text(
            family = report_font,
            face = "plain",
            size = 16,
            hjust = 0.5
          ),
          plot.subtitle = element_text(hjust = 0.5, color = "grey30", margin = margin(b = 10)),
          plot.margin = margin(14, 6, 14, 40)
        ) +
        labs(
          title = "Citing Policy Documents by Country/Region",
          subtitle = "Distinct policy documents in SciVal/Overton export; single-document regions grouped as Other"
        )
    )
}
```

---

# Media Opportunities

```{r media}
mediasumm <- metric_summary(
  metrics,
  category = "Media opportunity",
  value = "n",
  pre2019_category = "Media opportunity"
)

media_annual <- mediasumm %>%
  group_by(Year) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")

media_recent_years <- mediasumm %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

media_quarterly_recent <- mediasumm %>%
  filter(Year %in% media_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

media_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(media_annual$Year), max(media_annual$Year), by = 1)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("media")[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Number of Media Opportunities")

media_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("media")[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly Media Opportunities, Most Recent 5 Years")

media_type_quarterly <- metrics %>%
  filter(
    Category == "Media opportunity",
    !is.na(MediaOpportunityType),
    MediaOpportunityType != "",
    Year >= 2026
  ) %>%
  arrange(Year, Quarter) %>%
  mutate(
    MediaOpportunityType = fct_lump_min(MediaOpportunityType, min = 2, other_level = "Other single-entry"),
    YearQuarterLabel = if_else(
      Year == current_year & Quarter == current_quarter,
      paste(Year, Quarter, "(QTD)"),
      paste(Year, Quarter)
    ),
    YearQuarter = factor(YearQuarterLabel, levels = unique(YearQuarterLabel))
  ) %>%
  count(Year, Quarter, YearQuarter, MediaOpportunityType, name = "Quantity") %>%
  complete(
    YearQuarter,
    MediaOpportunityType,
    fill = list(Quantity = 0)
  ) %>%
  group_by(MediaOpportunityType) %>%
  mutate(TypeTotal = sum(Quantity, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(MediaOpportunityType = fct_reorder(MediaOpportunityType, TypeTotal))

ggplot(media_type_quarterly, aes(x = YearQuarter, y = MediaOpportunityType, fill = Quantity)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))), size = 3.5) +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11)
  ) +
  scale_fill_gradient(low = scales::alpha(metric_colors[["media"]], 0.1), high = metric_colors[["media"]]) +
  labs(
    x = "",
    y = "",
    title = "Quarterly Media Opportunity Type Breakdown",
    subtitle = "Rows with media type detail, starting in 2026"
  )
```

---

# Presentations

```{r presentations}
pressumm <- metric_summary(
  metrics,
  category = "Presentation",
  value = "n",
  pre2019_category = "Presentation"
)

presentation_annual <- pressumm %>%
  group_by(Year) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")

presentation_recent_years <- pressumm %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

presentation_quarterly_recent <- pressumm %>%
  filter(Year %in% presentation_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

presentation_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(presentation_annual$Year), max(presentation_annual$Year), by = 1)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("presentations")[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Number of Presentations")

presentation_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("presentations")[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly Presentations, Most Recent 5 Years")

presentation_type_quarterly <- metrics %>%
  filter(
    Category == "Presentation",
    !is.na(PresentationType),
    PresentationType != "",
    Year >= 2026
  ) %>%
  arrange(Year, Quarter) %>%
  mutate(
    PresentationType = fct_lump_min(PresentationType, min = 2, other_level = "Other single-entry"),
    YearQuarterLabel = if_else(
      Year == current_year & Quarter == current_quarter,
      paste(Year, Quarter, "(QTD)"),
      paste(Year, Quarter)
    ),
    YearQuarter = factor(YearQuarterLabel, levels = unique(YearQuarterLabel))
  ) %>%
  count(Year, Quarter, YearQuarter, PresentationType, name = "Quantity") %>%
  complete(
    YearQuarter,
    PresentationType,
    fill = list(Quantity = 0)
  ) %>%
  group_by(PresentationType) %>%
  mutate(TypeTotal = sum(Quantity, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(PresentationType = fct_reorder(PresentationType, TypeTotal))

ggplot(presentation_type_quarterly, aes(x = YearQuarter, y = PresentationType, fill = Quantity)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))), size = 3.5) +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11)
  ) +
  scale_fill_gradient(low = scales::alpha(metric_colors[["presentations"]], 0.1), high = metric_colors[["presentations"]]) +
  labs(
    x = "",
    y = "",
    title = "Quarterly Presentation Type Breakdown",
    subtitle = "Rows with presentation type detail, starting in 2026"
  )
```

---

# Total Dissemination Outputs

```{r dissemination}
dissem <- bind_rows(
  `Scientific publications` = pubsumm,
  Media = mediasumm,
  Presentations = pressumm,
  .id = "Category"
)

dissem_annual <- dissem %>%
  filter(Year >= 2017) %>%
  group_by(Year, Category) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    Category = factor(Category, levels = c("Presentations", "Media", "Scientific publications")),
    YearLabel = if_else(Year == current_year, paste0(Year, " (YTD)"), as.character(Year)),
    YearLabel = factor(YearLabel, levels = unique(YearLabel[order(Year)]))
  )

dissem_totals <- dissem_annual %>%
  group_by(Year, YearLabel) %>%
  summarise(Total = sum(Quantity, na.rm = TRUE), .groups = "drop")

ggplot(dissem_annual, aes(x = YearLabel, y = Quantity, fill = Category)) +
  geom_col(alpha = 0.85) +
  geom_text(
    data = dissem_totals,
    aes(x = YearLabel, y = Total, label = scales::comma(Total)),
    inherit.aes = FALSE,
    vjust = -0.35,
    color = "black",
    size = 3
  ) +
  theme_custom() +
  theme(axis.text.x = element_text(size = 10), legend.position = "bottom") +
  scale_fill_manual(values = c(
    "Presentations" = metric_colors[["presentations"]],
    "Media" = metric_colors[["media"]],
    "Scientific publications" = metric_colors[["publications"]]
  )) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(x = "", y = "", title = "Number of Dissemination Outputs")
```

---

# Capacity Sharing & Training

```{r capacity-sharing}
engagement_unique_ind <- metric_summary(
  metrics,
  category = "Capacity Sharing & Training",
  value = "quantity",
  pre2019_category = "Citizen/stakeholder engagement"
)

capacity_annual <- engagement_unique_ind %>%
  group_by(Year) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")

capacity_recent_years <- engagement_unique_ind %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

capacity_quarterly_recent <- engagement_unique_ind %>%
  filter(Year %in% capacity_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

capacity_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(capacity_annual$Year), max(capacity_annual$Year), by = 1)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("capacity")[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "People Engaged Through Capacity Sharing & Training")

capacity_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("capacity")[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly Capacity Sharing & Training Engagement, Most Recent 5 Years")

capacity_type_quarterly <- metrics %>%
  filter(
    Category == "Capacity Sharing & Training",
    !is.na(CapacitySharingType),
    CapacitySharingType != "",
    Year >= 2026
  ) %>%
  arrange(Year, Quarter) %>%
  mutate(
    YearQuarterLabel = if_else(
      Year == current_year & Quarter == current_quarter,
      paste(Year, Quarter, "(QTD)"),
      paste(Year, Quarter)
    ),
    YearQuarter = factor(YearQuarterLabel, levels = unique(YearQuarterLabel))
  ) %>%
  group_by(YearQuarter, CapacitySharingType) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop") %>%
  complete(
    YearQuarter,
    CapacitySharingType,
    fill = list(Quantity = 0)
  ) %>%
  group_by(CapacitySharingType) %>%
  mutate(TypeTotal = sum(Quantity, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(CapacitySharingType = fct_reorder(CapacitySharingType, TypeTotal))

ggplot(capacity_type_quarterly, aes(x = YearQuarter, y = CapacitySharingType, fill = Quantity)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(Quantity == 0, "", scales::comma(Quantity))), size = 3.5) +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 9),
    axis.text.y = element_text(size = 11)
  ) +
  scale_fill_gradient(low = scales::alpha(metric_colors[["capacity"]], 0.1), high = metric_colors[["capacity"]]) +
  labs(
    x = "",
    y = "",
    title = "Quarterly Capacity Sharing & Training Type Breakdown",
    subtitle = "Rows with capacity sharing & training type detail, starting in 2026"
  )
```

---

# Field Research

```{r field-research}
field_days <- metrics %>%
  filter(Category == "Field research") %>%
  group_by(Year, Quarter) %>%
  summarise(Quantity = sum(FieldDays, na.rm = TRUE), .groups = "drop") %>%
  bind_rows(filter(pre2019, Category == "Field research") %>% select(Year, Quarter, Quantity)) %>%
  mutate(Quarter = factor(Quarter, levels = quarter_levels)) %>%
  complete(
    Year = full_seq(range(Year, na.rm = TRUE), 1),
    Quarter = factor(quarter_levels, levels = quarter_levels),
    fill = list(Quantity = 0)
  )

field_annual <- field_days %>%
  group_by(Year) %>%
  summarise(Quantity = sum(Quantity, na.rm = TRUE), .groups = "drop")

field_by_department <- metrics %>%
  filter(Category == "Field research", Year >= 2026) %>%
  select(
    Year,
    Quarter,
    `Conservation Research` = FieldConservationResearch,
    `Animal Health` = FieldAnimalHealth,
    `Animal Care` = FieldAnimalCare,
    `Other Shedd departments` = FieldOtherDepartments
  ) %>%
  pivot_longer(
    cols = -c(Year, Quarter),
    names_to = "Department",
    values_to = "PersonDays"
  ) %>%
  filter(!is.na(PersonDays), PersonDays > 0) %>%
  arrange(Year, Quarter) %>%
  mutate(
    YearQuarterLabel = if_else(
      Year == current_year & Quarter == current_quarter,
      paste(Year, Quarter, "(QTD)"),
      paste(Year, Quarter)
    ),
    YearQuarter = factor(YearQuarterLabel, levels = unique(YearQuarterLabel))
  ) %>%
  group_by(YearQuarter, Department) %>%
  summarise(PersonDays = sum(PersonDays, na.rm = TRUE), .groups = "drop") %>%
  complete(
    YearQuarter,
    Department,
    fill = list(PersonDays = 0)
  ) %>%
  group_by(Department) %>%
  mutate(DepartmentTotal = sum(PersonDays, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Department = fct_reorder(Department, DepartmentTotal))

field_recent_years <- field_days %>%
  distinct(Year) %>%
  arrange(desc(Year)) %>%
  slice_head(n = 5) %>%
  pull(Year)

field_quarterly_recent <- field_days %>%
  filter(Year %in% field_recent_years) %>%
  mutate(PeriodStatus = if_else(Year == current_year & Quarter == current_quarter, "Quarter to date", "Complete period"))

field_annual %>%
  mutate(PeriodStatus = if_else(Year == current_year, "Year to date", "Complete period")) %>%
  ggplot(aes(x = Year, y = Quantity)) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(round(Quantity, 0)))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = seq(min(field_annual$Year), max(field_annual$Year), by = 1)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("field")[c("Complete period", "Year to date")]) +
  labs(x = "", y = "", title = "Field Research Days")

field_quarterly_recent %>%
  ggplot(aes(x = Quarter, y = Quantity)) +
  facet_grid(~ Year) +
  geom_col(aes(fill = PeriodStatus), alpha = 0.8) +
  geom_text(
    aes(label = ifelse(Quantity == 0, "", scales::comma(round(Quantity, 0)))),
    color = "black",
    vjust = -0.35,
    size = 3
  ) +
  theme_custom() +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = period_fill("field")[c("Complete period", "Quarter to date")]) +
  labs(x = "", y = "", title = "Quarterly Field Research Days, Most Recent 5 Years")

ggplot(field_by_department, aes(x = YearQuarter, y = Department, fill = PersonDays)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(PersonDays == 0, "", scales::comma(round(PersonDays, 0)))), size = 3.5) +
  theme_custom() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11)
  ) +
  scale_fill_gradient(low = scales::alpha(metric_colors[["field"]], 0.1), high = metric_colors[["field"]]) +
  labs(
    x = "",
    y = "",
    title = "Quarterly Field Person-Days by Shedd Department",
    subtitle = "Rows with field team detail, starting in 2026"
  )
```

---

# Description Text Themes

```{r description-publication-title-word-cloud, fig.height=6, fig.width=10}
stop_words_custom <- c(
  "a", "about", "above", "across", "after", "again", "against", "all", "also", "am", "an", "and",
  "any", "are", "as", "at", "be", "because", "been", "being", "below", "between", "both", "but",
  "by", "can", "could", "did", "do", "does", "doing", "down", "during", "each", "few", "for",
  "from", "further", "had", "has", "have", "having", "he", "her", "here", "hers", "him", "his",
  "how", "i", "if", "in", "into", "is", "it", "its", "just", "me", "more", "most", "my", "no",
  "nor", "not", "of", "off", "on", "once", "only", "or", "other", "our", "out", "over", "own",
  "same", "she", "should", "so", "some", "such", "than", "that", "the", "their", "them", "then",
  "there", "these", "they", "this", "those", "through", "to", "too", "under", "until", "up",
  "very", "was", "we", "were", "what", "when", "where", "which", "while", "who", "whom", "why",
  "will", "with", "would", "you", "your",
  "shedd", "aquarium", "research", "conservation", "project", "program", "team", "staff",
  "department", "departments", "work", "worked", "working", "data", "using", "including",
  "included", "provide", "provided"
)

tokenize_terms <- function(text, max_n = 3) {
  tokens <- text %>%
    str_to_lower() %>%
    str_replace_all("[^a-z0-9\\s-]", " ") %>%
    str_replace_all("\\b[0-9]+\\b", " ") %>%
    str_split("\\s+") %>%
    unlist() %>%
    str_replace_all("^-+|-+$", "")

  tokens <- tokens[
    tokens != "" &
      nchar(tokens) >= 3 &
      !tokens %in% stop_words_custom
  ]

  map_dfr(seq_len(max_n), function(n) {
    if (length(tokens) < n) {
      return(tibble())
    }

    tibble(
      Term = map_chr(seq_len(length(tokens) - n + 1), ~ paste(tokens[.x:(.x + n - 1)], collapse = " ")),
      Words = n
    )
  })
}

extract_publication_title <- function(citation) {
  citation <- str_squish(as.character(citation))
  citation <- str_replace_all(citation, "https?://[^ ]+|doi:? *[^ ]+", "")

  title <- case_when(
    str_detect(citation, "\\([12][0-9]{3}[^)]*\\)\\.?\\s+") ~
      str_match(citation, "\\([12][0-9]{3}[^)]*\\)\\.?\\s+(.+?)(?:\\.\\s+[A-Z][A-Za-z& ,:-]{2,}|$)")[, 2],
    str_detect(citation, "\\b[12][0-9]{3}\\.\\s+") ~
      str_match(citation, "\\b[12][0-9]{3}\\.\\s+(.+?)(?:\\.\\s+[A-Z][A-Za-z& ,:-]{2,}|$)")[, 2],
    TRUE ~
      str_match(citation, "^(.+?)(?:\\.\\s+[A-Z][A-Za-z& ,:-]{2,}|$)")[, 2]
  )

  title <- coalesce(title, citation)
  title <- str_squish(str_replace_all(title, "^[[:punct:] ]+|[[:punct:] ]+$", ""))
  if_else(nchar(title) >= 8, title, citation)
}

description_documents <- bind_rows(
  metrics %>%
    select(Year, Date, Category, Description),
  metrics_2022_2024_raw %>%
    transmute(
      Date = first_non_missing_date(
        as.Date(`Publication Date`),
        as.Date(`Media Opportunity Date`),
        as.Date(`What was the date of the engagement?`),
        as.Date(`What was the starting date of the engagement?`),
        as.Date(`Start time`)
      ),
      Year = year(Date),
      Category = standardize_category(`What was the engagement type?`),
      Description = `Please provide a description of this media engagement`
    )
) %>%
  mutate(Description = str_squish(as.character(Description))) %>%
  filter(!is.na(Description), Description != "") %>%
  distinct(Year, Date, Category, Description)

publication_title_documents <- metrics %>%
  filter(Category == "Publication", !is.na(Citation), Citation != "") %>%
  transmute(
    Year,
    Date,
    Category = "Publication title",
    Description = extract_publication_title(Citation)
  ) %>%
  mutate(Description = str_squish(as.character(Description))) %>%
  filter(
    !is.na(Description),
    Description != "",
    !str_detect(Description, "^[A-Z][A-Za-z'’.-]+,\\s*[A-Z]"),
    !str_detect(Description, "^[A-Z][A-Za-z'’.-]+\\s+[A-Z]{1,3}\\.?,"),
    !str_detect(Description, "\\([12][0-9]{3}$")
  ) %>%
  distinct(Description, .keep_all = TRUE)

description_publication_terms <- bind_rows(
  description_documents,
  publication_title_documents
) %>%
  mutate(Terms = map(Description, tokenize_terms)) %>%
  unnest(Terms)

description_publication_term_counts <- description_publication_terms %>%
  count(Term, Words, name = "Entries") %>%
  filter(
    (Words == 1 & Entries >= 8) |
      (Words == 2 & Entries >= 3) |
      (Words == 3 & Entries >= 2)
  ) %>%
  filter(str_count(Term, "[a-z0-9]") >= 3) %>%
  mutate(Score = Entries * (1 + 0.35 * (Words - 1)))

set.seed(2026)
description_publication_word_cloud_terms <- description_publication_term_counts %>%
  slice_max(Score, n = 80, with_ties = FALSE) %>%
  mutate(
    LabelSize = scales::rescale(sqrt(Score), to = c(4, 16)),
    TermType = case_when(
      Words == 1 ~ "Word",
      Words == 2 ~ "Two-word phrase",
      TRUE ~ "Three-word phrase"
    ),
    FrequencyTier = ntile(desc(Score), 5)
  ) %>%
  group_by(FrequencyTier) %>%
  slice_sample(prop = 1) %>%
  ungroup() %>%
  arrange(FrequencyTier)

description_publication_word_cloud_terms %>%
  ggplot(aes(label = Term, size = LabelSize, color = TermType)) +
  ggwordcloud::geom_text_wordcloud(
    seed = 2026,
    eccentricity = 0.75,
    grid_margin = 3,
    max_steps = 80,
    rm_outside = FALSE
  ) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 20, 10, 20)
  ) +
  scale_size_identity() +
  scale_color_manual(values = c("Word" = "#4D4D4D", "Two-word phrase" = "#2C7FB8", "Three-word phrase" = "#D95F02")) +
  labs(
    title = "Common Description and Publication Title Words and Phrases",
    subtitle = "Description text plus titles parsed from publication citations; size uses square-root scaled frequency",
    color = ""
  )
```

---

# Data Quality Checks

```{r data-quality}
data_quality_summary <- tibble(
  Check = c(
    "Rows missing category or date",
    "Duplicate-looking rows removed"
  ),
  Count = c(
    nrow(metrics_all %>% filter(is.na(Category) | is.na(Date))),
    nrow(metrics_duplicates_removed)
  )
)

data_quality_summary %>%
  knitr::kable(caption = "Data quality summary")

rows_needing_review <- metrics_all %>%
  filter(is.na(Category) | is.na(Date)) %>%
  select(Source, ID, Date, RawCategory, Category, Description)

if (nrow(rows_needing_review) > 0) {
  rows_needing_review %>%
  knitr::kable(caption = "Rows needing category or date review")
} else {
  tibble(Status = "No retained rows are missing category or date.") %>%
    knitr::kable(caption = "Rows needing category or date review")
}

if (nrow(metrics_duplicates_removed) > 0) {
  metrics_duplicates_removed %>%
    count(Source, Category, sort = TRUE, name = "RowsRemoved") %>%
    knitr::kable(caption = "Duplicate-looking rows removed by source and category")
} else {
  tibble(Status = "No duplicate-looking rows were removed.") %>%
    knitr::kable(caption = "Duplicate-looking rows removed by source and category")
}
```

## Possible Outlier Review

```{r possible-outlier-review}
robust_upper_limit <- function(x, floor = 0) {
  x <- x[!is.na(x) & is.finite(x) & x > 0]
  if (length(x) < 4) {
    return(Inf)
  }

  max(floor, quantile(x, 0.75, names = FALSE) + 3 * IQR(x))
}

outlier_limits <- metrics %>%
  group_by(Category) %>%
  summarise(
    QuantityLimit = robust_upper_limit(Quantity, floor = 10),
    FieldDaysLimit = robust_upper_limit(FieldDays, floor = 30),
    FieldPersonDaysLimit = robust_upper_limit(FieldPersonDays, floor = 40),
    .groups = "drop"
  )

possible_outliers <- metrics %>%
  left_join(outlier_limits, by = "Category") %>%
  mutate(
    QuantityFlag = Category != "Presentation" & !is.na(Quantity) & Quantity > QuantityLimit,
    FieldDaysFlag = Category == "Field research" & !is.na(FieldDays) & FieldDays > FieldDaysLimit,
    FieldPersonDaysFlag = Category == "Field research" & !is.na(FieldPersonDays) & FieldPersonDays > FieldPersonDaysLimit,
    FlagReason = pmap_chr(
      list(QuantityFlag, FieldDaysFlag, FieldPersonDaysFlag, Quantity, QuantityLimit, FieldDays, FieldDaysLimit, FieldPersonDays, FieldPersonDaysLimit),
      function(quantity_flag, field_days_flag, field_person_days_flag, quantity, quantity_limit, field_days, field_days_limit, field_person_days, field_person_days_limit) {
        reasons <- c()

        if (quantity_flag) {
          reasons <- c(reasons, paste0("Quantity ", scales::comma(quantity), " exceeds category threshold ", scales::comma(round(quantity_limit, 1))))
        }
        if (field_days_flag) {
          reasons <- c(reasons, paste0("Field days ", scales::comma(field_days), " exceeds field threshold ", scales::comma(round(field_days_limit, 1))))
        }
        if (field_person_days_flag) {
          reasons <- c(reasons, paste0("Field person-days ", scales::comma(field_person_days), " exceeds field threshold ", scales::comma(round(field_person_days_limit, 1))))
        }

        paste(reasons, collapse = "; ")
      }
    )
  ) %>%
  filter(FlagReason != "") %>%
  transmute(
    Source,
    ID,
    Date,
    Category,
    Name,
    Quantity,
    FieldDays,
    FieldPersonDays,
    Description = str_trunc(Description, 90),
    FlagReason
  ) %>%
  arrange(Category, desc(Quantity), desc(FieldPersonDays), Date)

if (nrow(possible_outliers) > 0) {
  possible_outliers_display <- possible_outliers %>%
    select(where(~ !all(is.na(.x))))

  possible_outliers %>%
    select(names(possible_outliers_display)) %>%
    knitr::kable(caption = "Possible outlier rows needing manual review")
} else {
  tibble(Status = "No extreme within-category outliers were detected.") %>%
    knitr::kable(caption = "Possible outlier rows needing manual review")
}
```
